### 2021-04-06 img/kernel_header.tex
OpenCL highlighting

\begin{opencl}
#include "bempp_base_types.h"
#include "bempp_helpers.h"
#include "bempp_spaces.h"
#include "kernels.h"

__kernel void kernel_function(
    __global uint* testIndices, __global uint* trialIndices,
    __global int* testNormalSigns, __global int* trialNormalSigns,
    __global REALTYPE* testGrid, __global REALTYPE* trialGrid,
    __global uint* testConnectivity, __global uint* trialConnectivity,
    __global uint* testLocal2Global, __global uint* trialLocal2Global,
    __global REALTYPE* testLocalMultipliers, __global REALTYPE* trialLocalMultipliers,
    __constant REALTYPE* quadPoints, __constant REALTYPE* quadWeights,
    __global REALTYPE* globalResult,
    __global REALTYPE* kernel_parameters,
    int nTest, int nTrial, char gridsAreDisjoint)
{
  /* */
}
\end{opencl}