### 2021-04-09 numba.tex
Started performance section

\section{NUMBA ASSEMBLY OF INTEGRAL OPERATORS}

The main focus of Numba within Bempp-cl is to provide accelerated implementations for linear complexity routines, such as grid iterations, integration of functions over grids or assembly of certain sparse matrices. However, we do also provide a fall-back implementation of the OpenCL dense operators in Numba. Here, we use classic loop parallelism. Each loop iteration is the assembly of one test element with all trial elements. Within each loop we try to optimize for auto-vectorization by linearly passing through the data in memory order for the individual operations. However, a much smaller fraction of operations is SIMD optimized due to the lack of targeted SIMD constructs in Numba.