### 2021-04-14 main.tex
remove template

\documentclass{IEEEcsmag}

\usepackage[colorlinks,urlcolor=blue,linkcolor=blue,citecolor=blue]{hyperref}

\usepackage{xcolor}
\definecolor{bemppgrey}{HTML}{555555}
\definecolor{bempporange}{HTML}{FF7A1A}

\usepackage{pythonhighlight}

\usepackage{tikz}
\usetikzlibrary{arrows,decorations.pathreplacing}
\usepackage{ifthen}

\usepackage{upmath}
\usepackage{upgreek}
\usepackage{amsmath, amssymb}

\usepackage{cleveref}
\crefname{equation}{Equation}{Equations}
\Crefname{equation}{Equation}{Equations}
\crefname{figure}{Figure}{Figures}
\Crefname{figure}{Figure}{Figures}
\crefname{section}{Section}{Sections}
\Crefname{section}{Section}{Sections}

% Bold vector (for points in R^3)
\newcommand{\bvec}[1]{\boldsymbol{#1}}
\def\bx{\bvec{x}}
\def\by{\bvec{y}}
\newcommand{\ds}[1][]{\,\mathrm{d}s\ifthenelse{\equal{#1}{}}{}{_{#1}}}

% Discrete vector and discrete matrix
\newcommand{\dmat}[1]{\mathbf{#1}}
\newcommand{\dvec}[1]{\mathbf{#1}}

\newcommand{\ii}{\mathrm{i}}
\newcommand{\ee}{\mathrm{e}}

\jvol{XX}
\jnum{XX}
\paper{8}
\jmonth{May/June}
\jname{IT Professional}
\pubyear{2021}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}

\setcounter{secnumdepth}{0}

\begin{document}

\sptitle{Department: Head}
\editor{Editor: Name, xxxx@email}

\title{Designing a high-performance boundary element library with OpenCL and Numba}

\author{T. Betcke}
\affil{Department of Mathematics, University College London}

\author{M. W. Scroggs}
\affil{Department of Engineering, University of Cambridge}

\markboth{Department Head}{Paper title}

\begin{abstract}
The Bempp boundary element library is a well known library for the simulation of a range of electrostatic, acoustic and electromagnetic problems in homogeneous bounded and unbounded domains. It originally started as a traditional C++ library with a Python interface. Over the last two years we have completely redesigned Bempp as a native Python library, called Bempp-cl, that provides computational backends for OpenCL (using PyOpenCL) and Numba. The OpenCL backend implements kernels for GPUs and SIMD optimized for CPUs. In this paper we will discuss the design of Bempp-cl, provide performance comparisons on different compute devices and discuss the advantages and disadvantages of OpenCL as compared to Numba.
\end{abstract}

\maketitle

\input{intro}
\input{bem}
\input{bempp_overview}
\input{opencl}
\input{numba}
\input{performance}

\bibliographystyle{IEEEtran}
\begin{thebibliography}{1}

\bibitem{Bempp-cl}
T. Betcke and M. W. Scroggs,
``Bempp-cl: A fast Python based just-in-time compiling boundary element library,''
{\it Journal of Open Source Software}, vol. 6, no. 59, p. 2879, 2021.
\textsc{doi}: \href{https://dx.doi.org/10.21105/joss.02879}{10.21105/joss.02879}.

\bibitem{pyopencl}
A. Kloeckner, N. Pinto, Y. Lee, B. Catanzaro, P. Ivanov, and A. Fasih,
``PyCUDA and PyOpenCL: A scripting-based approach to GPU run-time code generation,''
{\it Parallel Computing}, vol. 38, no. 3, pp. 157--174, 2012.
\textsc{doi}: \href{https://dx.doi.org/10.1016/j.parco.2011.09.001}{10.1016/j.parco.2011.09.001}.

\bibitem{opencl}
OpenCL, \url{https://www.khronos.org/opencl/}.

\bibitem{bempp_maxwell}
M. W. Scroggs, T. Betcke, E. Burman, Wojciech \'{S}migaj, and E. van 't Wout, ``Software frameworks for integral equations in electromagnetic scattering based on Cader\'{o}n identities''
{\it Computers \& Mathematics with Applications}, vol. 74, no. 11, pp. 2897--2914, 2017.
\textsc{doi}: \href{https://doi.org/10.1016/j.camwa.2017.07.049}{10.1016/j.camwa.2017.07.049}.

\bibitem{bempp_orig}
W. \'{S}migaj, T. Betcke, S. Arridge, J. Phillips, and M. Schweiger,
``Solving boundary integral problems with BEM++,''
{\it ADM Transactions on Mathematical Software}, vol. 41, no. 2, Article 6, 2015.
\textsc{doi}: \href{https://doi.org/10.1145/2590830}{10.1145/2590830}

\bibitem{bempp_exafmm}
T. Wang, C. D. Cooper, T. Betcke, and L. A. Barba, ``High-productivity, high-performance workflow for virus-scale electrostatic simulations with Bempp-Exafmm''
{\it arXiv preprint} arXiv:103.01048, 2021.
\end{thebibliography}

\end{document}